---
title: "analysis"
author: "Dan Birman"
date: "Thursday, February 19, 2015"
output: html_document
---

```{r}
library(jsonlite)
library(ggplot2)
library(grid)
setwd('C:\\Users\\Dan\\proj\\repo254\\production-results\\')
# setwd('/Users/dan/proj/repo254/production-results/')
```

```{r}

parseData = function(file) {
#   print(file)
  print(data$answer$fingerprint$ip)
  data = fromJSON(file)
  if (nrow(data$answer$trialInfo) > 15) {
    print('This person left fullscreen...')
    if (file=='3EG49X351UCD25R1OOP3Z039Z03X69.json') {
      data$answer$trialInfo = data$answer$trialInfo[-c(15,16,17,18), ]
      data$answer$t_stream = data$answer$t_stream[-c(5,6,7,8),]
    }
  }
    
  # pull out the 'answer' variable
  ans = data$answer
  
#   print(ans$demo$age)
#   print(ans$demo$sex)
  # get the big picture data
  fingerprint = ans$fingerprint
  # now get the trial data
  training = ans$t_train
  critical = ans$t_critical
  backg = ans$t_backg
  stream = ans$t_stream
  trials = ans$trialInfo
  
  if (trials$backgtask[6]==1) {
    backg_pos = c(6:10)
    notbackg = c(11:15)
  } else {
    backg_pos = c(11:15)
    notbackg = c(6:10)
  }
  trials$correct[1:15] = 0
  trials$correct[1:4] = training$regResp == trials$digits[1:4]
  trials$correct[notbackg] = stream$regResp == trials$digits[notbackg]
  trials$correct[5] = critical$catchImageResp == trials$catchImage[5]
  trials$correct[backg_pos] = backg$catchImageResp == trials$catchImage[backg_pos]
  backg_immed = 0
  backg_immed_inc = 0
  backg_class = 0
  backg_blind = 0
  for (i in 1:5) {
    t = backg_pos[i]
    if (backg$catchResp1[i]=='Y') {
      if (trials$correct[t]==1) {
        backg_immed = backg_immed + 1
      } else {
        backg_immed_inc = backg_immed_inc + 1
      }
    } else {
      if (trials$correct[t]==1) {
        backg_class = backg_class + 1
      } else {
        backg_blind = backg_blind + 1
      }
    }
  }
  
  train_perf = mean(trials$correct[1:4])
  if (critical$catchResp1=='Y') {
    if (trials$correct[5]==1) {
      crit_perf = 'immediate'
    } else {
      crit_perf = 'immediate_inc'
    }
  } else {
    if (trials$correct[5]==1) {
      crit_perf = 'classified'
    } else {
      crit_perf = 'blind'
    }
  }
  backg_perf = mean(trials$correct[backg_pos])
  stream_perf = mean(trials$correct[notbackg])

  return(data.frame(train_perf,crit_perf,backg_perf,stream_perf,backg_immed,backg_immed_inc,backg_class,backg_blind,catchImg=trials$catchImage[5],catchCat=substr(trials$catchImage[5],1,1)))
}
```

```

Get Files

```{r}
infiles <- dir(pattern='.json$')

out_rs = data.frame()
# skipping: 4
for (i in c(1:3,5:30)) {
  out_rs = rbind(out_rs,parseData(infiles[i]))
}
```

```{r}
t.test2 <- function(m1,m2,s1,s2,n1,n2,m0=0,equal.variance=FALSE)
{
    if( equal.variance==FALSE ) 
    {
        se <- sqrt( (s1^2/n1) + (s2^2/n2) )
        # welch-satterthwaite df
        df <- ( (s1^2/n1 + s2^2/n2)^2 )/( (s1^2/n1)^2/(n1-1) + (s2^2/n2)^2/(n2-1) )
    } else
    {
        # pooled standard deviation, scaled by the sample sizes
        se <- sqrt( (1/n1 + 1/n2) * ((n1-1)*s1^2 + (n2-1)*s2^2)/(n1+n2-2) ) 
        df <- n1+n2-2
    }      
    t <- (m1-m2-m0)/se 
    dat <- c(m1-m2, se, t, 2*pt(-abs(t),df))    
    names(dat) <- c("Difference of means", "Std Error", "t", "p-value")
    return(dat) 
}
```

Results

```{r}
blind_perc = length(out_rs$crit_perf[out_rs$crit_perf=='blind']) / nrow(out_rs)
class_perc = length(out_rs$crit_perf[out_rs$crit_perf=='classified']) / nrow(out_rs)
immed_perc = length(out_rs$crit_perf[out_rs$crit_perf=='immediate']) / nrow(out_rs)
immed_inc = length(out_rs$crit_perf[out_rs$crit_perf=='immediate_inc']) / nrow(out_rs)

backg_blind_perc = mean(out_rs$backg_blind/5)
backg_class_perc = mean(out_rs$backg_class/5)
backg_immed_perc = mean(out_rs$backg_immed/5)
backg_immed_inc_perc = mean(out_rs$backg_immed_inc/5)

backg_list = mean(out_rs$backg_perf)
train_perf = mean(out_rs$train_perf)
train_sd = sd(out_rs[,1])/sqrt(nrow(out_rs))
stream_perf = mean(out_rs[,4])
stream_sd = sd(out_rs[,4])/sqrt(nrow(out_rs))
```

```{r}
values = c(0.5,blind_perc+immed_inc,backg_immed_inc_perc+backg_blind_perc,0.23,immed_perc,backg_immed_perc,1-.5-.23,class_perc,backg_class_perc)
ses = sqrt((values*(1-values))/c(30,29,29,30,29,29,30,29,29))

out = data.frame(value = values, sd = ses, dataset=c(0,1,2,0,1,2,0,1,2),
                 group=c(1,1,1,2,2,2,3,3,3))
out$dataset = factor(out$dataset,levels=c(0,1,2),labels=c("Cohen et al. 2011","Birman Replication","Birman Rep: Attending\nto Background"))
out$group = factor(out$group,levels=c(1,2,3),labels=c("Blind","Chose Correct Image\nImmediately","Chose Correct Image\nAfter Questions"))
ggplot(out,aes(group,value,fill=dataset)) +
  geom_bar(stat="identity",position="dodge",width=.75) +
  geom_errorbar(aes(ymin=value-sd,ymax=value+sd),position="dodge",width=.75,color='darkgray') +
  xlab('') +
  ylab('% of Participants') +
  ggtitle("Results: Cohen et al. 2011 (n=30) -- Birman Replication (n=29)") +
  theme_bw() +
  scale_y_continuous(limits=c(0,1)) +
  scale_fill_brewer(palette='YlOrBr') +
  theme(panel.border=element_blank(),axis.ticks.x=element_blank())
```

```{r}
values = c(blind_perc,backg_blind_perc,immed_inc,backg_immed_inc_perc,immed_perc,backg_immed_perc,class_perc,backg_class_perc)
ses = sqrt((values*(1-values))/c(29,29,29,29,29,29,29,29))

out = data.frame(value = values, sd = ses, dataset=c(1,2,1,2,1,2,1,2),
                 group=c(0,0,1,1,2,2,3,3))
out$dataset = factor(out$dataset,levels=c(1,2),labels=c("Birman Replication","Birman Rep: Attending\nto Background"))
out$group = factor(out$group,levels=c(0,1,2,3),labels=c("Blind","Saw ''Something''\nBut Chose Wrong Image", "Saw Image Immediately","Classified After Questions"))
ggplot(out,aes(group,value,fill=dataset)) +
  geom_bar(stat="identity",position="dodge",width=.75) +
  geom_errorbar(aes(ymin=value-sd,ymax=value+sd),position="dodge",width=.75,color='darkgray') +
  xlab('') +
  ylab('% of Participants') +
  ggtitle("Results: Cohen et al. 2011 (n=30) -- Birman Replication (n=29)") +
  theme_bw() +
  scale_y_continuous(limits=c(0,1)) +
  scale_fill_brewer(palette='YlOrBr') +
  theme(panel.border=element_blank(),axis.ticks.x=element_blank())
```



```{r}
out = data.frame(value = c(.93,backg_immed_perc,.74,train_perf,.76,stream_perf),sd=c(NaN,sd((out_rs$backg_immed/5))/sqrt(29),.036,train_sd,.028,stream_sd),dataset=c(0,1,0,1,0,1),group=c(-1,-1,0,0,1,1))
out$dataset = factor(out$dataset,levels=c(0,1),labels=c("Cohen et al. 2011","Birman Replication"))
out$group = factor(out$group,levels=c(-1,0,1),labels=c("Background Task", "RSVP Task\nNo Background Image", "RSVP Task\nBackground Image"))

p1 = t.test2(out$value[1],out$value[2],out$sd[1],out$sd[2],8,29)
p2 = t.test2(out$value[3],out$value[4],out$sd[3],out$sd[4],8,29)

ggplot(out,aes(group,value,fill=dataset)) +
  geom_bar(stat="identity",position="dodge",width=.75) +
  geom_errorbar(aes(ymin=value-sd,ymax=value+sd),position="dodge",width=.75,color='darkgray') +
  scale_fill_brewer(palette='YlOrBr') +
  xlab('') +
  ylab("(% Correct)") +
  ggtitle("Control Task Performances: Cohen et al. 2011 Control (n=30,8,8) -- Birman Replication (n=29,29,29)") +
  scale_y_continuous(limits=c(0,1)) +
  theme_bw() +
  theme(panel.border=element_blank())
```


